---
title: "DSC640_Exercise_1-2_JakeMeyer"
course: "DSC 640 Data Presentation & Visualization (DSC640-T302 2231-1)"
college: "Bellevue University"
author: "Jake Meyer"
date: "2022-12-06"
output: pdf_document
always_allow_html: yes
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, 
                      root.dir = "C:/Users/jkmey/Documents/Github/DSC640_Course_Assignments/DSC640_Repository" )
```

## Assignment Instructions
Submit 1 bar chart, 1 stacked bar chart, 1 pie chart, and 1 donut chart with R.

### Show Working directory
```{r, warning = FALSE}
getwd()
dir()
 # setwd("C:/Users/jkmey/Documents/Github/DSC640_Course_Assignments/DSC640_Repository")
```

### Import the necessary libraries
```{r, warning = FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plotly)
theme_set(theme_minimal())
```

### Import the data
```{r}
df1 <- read_excel("hotdog-contest-winners.xlsm")
df2 <- read_excel("hotdog-places.xlsm")
df3 <- read_excel("obama-approval-ratings.xls")
```

### Revise data for Bar Charts
```{r}
df_dogs_eaten <- aggregate(df1$`Dogs eaten`, list(df1$Country), FUN = sum)
colnames(df_dogs_eaten) <- c('Country', 'DogsEaten')
df_dogs_eaten$Country <- factor(df_dogs_eaten$Country,
                                levels = df_dogs_eaten$Country[order(df_dogs_eaten$DogsEaten, decreasing = TRUE)])
```

### Bar Chart
```{r}
ggplot(df_dogs_eaten, aes(x = Country, y = DogsEaten, fill = Country, color = Country)) + 
  geom_bar(stat ='identity') + xlab('Country') + ylab('Count') + 
  ggtitle('Number of Hot Dogs Eaten by Country')
```

### Store data as specific variables
```{r}
x = df1$Country
y = df1$`Dogs eaten`
df1$`New record`=as.factor(df1$`New record`)
group = df1$`New record`
```

### Stacked Bar Chart for Hot Dog Eating
```{r}
ggplot(df1, aes(x = group, y = y, fill = x)) + 
  geom_bar(position = 'stack', stat='identity') + 
  xlab('New Record (0-No, 1-Yes)') + ylab('Number of Hot Dogs Eaten') + 
  ggtitle('Number of Hot Dogs Eaten by Country (New Record Breakdown)') +
  guides(fill = guide_legend(title = "Country")) 
```

### Stacked Bar Chart for Obama Approval Ratings Using Plotly
```{r}
fig <- plot_ly(df3, x = ~Issue, y = ~Approve, type = 'bar', name = 'Approve')
fig <- fig %>% add_trace(y = ~Disapprove, name = 'Disapprove')
fig <- fig %>% add_trace(y = ~None, name = 'None')
fig <- fig %>% layout(xaxis = list(cateogoryorder= "total descending"))
fig <- fig %>% layout(yaxis = list(title = 'Approval Percentage'), barmode = 'stack')
fig <- fig %>% layout(title = 'Obama Approval Ratings on Issues')
fig
```

### Revise data variables for Pie and Donut Charts
```{r}
percent_table <- table(c(df1$Country))
x_percent <- percent_table / length(df1$Country)
countries <- unique(df1$Country)
df_percent <- data.frame(x_percent)
colnames(df_percent) <- c('Countries', 'Percent')
y_new = df_percent$Percent
x_new = df_percent$Countries
```

### Pie Chart
```{r}
ggplot(df_percent, aes(x = "", y = y_new, fill = x_new)) + geom_col() + 
  coord_polar(theta = "y", start = 1) + 
  ggtitle('Percentage of Hot Dog Eating Champions by Country') +
  geom_text(aes(label = paste0(round(y_new*100), "%")),
             position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Country"))
```
### Donut Chart
```{r}
ggplot(df_percent, aes(x=2, y=y_new, fill = x_new)) + geom_col() + 
  coord_polar("y", start = 1) + 
  geom_text(aes(label = paste0(round(y_new*100), "%")),
            position = position_stack(vjust = 0.5)) +
  theme(panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size =18)) + 
  ggtitle("Percentage of Hot Dog Eating Champions by Country") + xlim(0.5, 2.5) +
  guides(fill = guide_legend(title = "Country"))
```

