---
title: "DSC640_Exercise_6-2_JakeMeyer"
course: "DSC 640 Data Presentation & Visualization (DSC640-T302 2231-1)"
college: "Bellevue University"
author: "Jake Meyer"
date: "2023-02-25"
output: pdf_document
always_allow_html: yes
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, 
                      root.dir = "C:/Users/jkmey/Documents/Github/DSC640_Course_Assignments/DSC640_Repository" )
```

## Assignment Instructions
Submit 1 histogram, 1 box plot, 1 bullet chart, and 1 additional chart with R.

### Show Working directory
```{r, warning = FALSE}
getwd()
dir()
 # setwd("C:/Users/jkmey/Documents/Github/DSC640_Course_Assignments/DSC640_Repository")
```

### Import the necessary libraries
```{r, warning = FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(scales)
library(plotly)
theme_set(theme_minimal())
```

### Import the data
```{r}
df1 <- read.csv("education.csv", header = TRUE, sep = ",")
df2 <- read.csv("education_revised.csv", header = TRUE, sep = ",")
```

### Histogram.
Generate a histogram using the education data.
```{r}
ggplot(df1, aes(writing)) + geom_histogram(binwidth = 7) +
  ggtitle('Writing Scores') + labs(x='Writing Score', y = 'Frequency') 
```


### Box Plot.
Generate a box plot using the education data.
```{r}
boxplot(df2$score~df2$category, ylab = 'Score', xlab = 'Subject',
        main =' US Math, Reading, Writing Scores')
```

### Bullet Chart.
Generate a bullet chart using the education data.
```{r}
ex_df <- bind_rows(
  tibble(
  name = rep("Reading", 2),
  group = c("Qualitative", "Measure"),
  color = c("grey", "black"),
  value = c(600, 536),
  width = c(0.9, 0.5),
  target = rep(580, 2),
  ymin = rep(0.7, 2),
  ymax = rep(1.3, 2)
  ),
  tibble(
  name = rep("Writing", 2),
  group = c("Qualitative", "Measure"),
  color = c("grey", "black"),
  value = c(600, 523),
  width = c(0.9, 0.5),
  target = rep(580, 2),
  ymin = rep(1.7, 2),
  ymax = rep(2.3, 2)
  )
)

ex_df %>% 
  ggplot(aes(x = value, y = name, fill = color)) +
  geom_col(width = c(0.9, 0.5, 0.9, 0.5)) +
  geom_linerange(
    aes(x = target, ymin = ymin, ymax = ymax),
    size = 2, color = "red"
    ) +
  coord_cartesian(ylim = c(0.3, 2.7)) +
  scale_fill_identity() +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank()) +
  labs(x = "Score", y = "Subject", title = "US Scores for Reading and Writing")

```

### Lollipop Chart.
Generate a lollipop chart with ggplot2.
```{r}
# Create the data for the chart.
data <- data.frame(
  x=LETTERS[1:26], y = abs(rnorm(26))
)

# Generate the plot.
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y), color="grey") +
  geom_point( color="green", size=5) +
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y") +
  ggtitle('26 Random Values - Lollipop Chart')
  


```